Title here

### 2017-01-03 portfolio/text/ms.Rmd
Develop discussion

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability.
A key component to managing these industries is identifying how regulatory
actions and individual behaviour affect income variability and by extension the
economic stability and sustainable use of natural resources. For commercial
fisheries, recent work has identified that communities and vessels can lower
revenue variability by increasing the diversity of species fished. 
However, it is unclear if ... extend to individual... year to year changes
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities and
investigate how individual actions along each axis affect revenue and revenue
variability. 
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. 
Our results demonstrate that the most effective option for individuals to
increase revenue and decrease variability is to purchase permits to participate
in additional or more diverse fisheries. However, this option is costly
and often limited by regulations such as limited entry and catch share
programs.
With increasing environmental variability in a changing climate, it will be
particularly important that individuals who rely on natural resources for
their livelihood have effective strategies to reduce income variability.

<!-- Thus, over the past 30 years as the Alaskan commercial fishing fleet has become -->
<!-- increasingly specialized, -->

@@

<!--This benefit of specializing is most apparent after accounting for fishing
effort ().-->

@@

Our findings across fishing strategies are in line with previous findings
[@kasperski2013, @sethi2012b, @sethi2014], but we also found that within
strategies, for individuals, specializing tended to be associated with increased
revenue and altering species diversity tended to be associated with increased
revenue variability. Specializing has been shown to be profitable in some
circumstances for other natural resource sectors. For example, while
diversification is generally helpful for small-scale farmers, farmers that
specialize on high-demand and high-value crops such as tobacco or cocoa can
benefit, at least in the short term, from specializing (REF). In terms of
variance, there are number of reasons by annual revenue variability might
increase with changes to individual fished species diversity. First, changes to
fishing patterns will sometimes be associated with a learning or adaptation
phase as individuals become accustomed to new locations, new gear, or new
species. Secondly, changes in diversity might reflect years in which targeted
species were less available and fishers were forced to target other species or
landed more bycatch. Most importantly, this increased variability technically
reflects an inability for our large-scale model to predict an individual's
revenue in a subsequent year. There are many reasons, often unique to an
individual permit holder and therefore not included in a large-scale model, that
might make that specific individual's revenue less predictable if species
diversity changes.

increased revenue with specialization? 
why might someone specialize? but this has long-term risks 
	efficiency 
		and individual behavior important in fleet dynamics and efficiency of catch @hilborn1985
		"Specialization, as measured by the share of total cultivated land devoted to rice production, also tends to lower inefficiency," @abdulai2000
	cash crops 
		sockeye e.g.
		cocoa: most profitable crop export in Nigeria @mustapha1999
			when cocoa prices were high, most cocoa farmers were highly specialized; now there is much more diversification to other sources of income... for example trading and electronic repairs @mustapha1999
		tobacco: high profitability and low variability of income from tobacco mean that the optimal farming strategy (regardless of risk profile) is to specialize as much as possible on tobacco (given alternatives such as soybean, alfalfa, chickpeas, ...) -- great reference to use -- quadratic optimization @chavez2014
	conversely, diversification can happen as a consequence of non-ideal situations 
		poor fishing years mean the fishermen must seek out other options - this paper 
		downwards economic shocks tend to be associated with increasing diversity for households relying on forests for income @debela2012
		for example, many small farms use diversification to maximize revenue given their limited land @barbieri2008
	but has long-term risks 
		"...most graphic, recent example of such risks was the reliance of U.S. crop producers on export markets for wheat, corn, and soybeans during the 1970s. Many farmers borrowed large sums of money to buy additional land and buy specialized equipment ....

		Specialized farmers producing export commodities were hardest hit by the financial crisis of American agriculture in the early 1980s...." @ikerd1990
		"Historically, cocoa farmers rarely engaged in non-farm commercial or artisanal activities. When cocoa farmers took to non-farm occupations in the 1970s and 19803, it was as a result of the severe crisis of accumulation confronting them (Berry 1985). Indeed, only 5 per cent of the 1964 Alade sample had any non-farm occupation (Oloko 1964: 57). My 1992 study established that by the early 1980s, that is just before SAP, about 35 per cent of the Lower stratum in that survey, 30 per cent of the Middle, and 33 percent of the Upper had non‐farm occupations." @mustapha1999

barriers to diversification 
	cost of entry
		"poorer households have fewer opportunities in non-cropping activities ... and hence less diversified incomes. ... reflect[s] their relative lack of capital, which makes it difficult for them to diversify away from subsistence agriculture...  " @abdulai2001
		entry barriers mean that poor farmers tend to be less diversified and are forced to experience lower returns and higher variability in earnings @barrett2001
			e.g. "Block and Webb and More income diversification associated with higher welfare measured in both income and nutritional terms in Ethiopia, and that the least diversified ft hard to increase their diversification over time. ... Barrett et al. report a strong association between greater income diversification ˆ and higher wealth and income in Cote d’Ivoire and Kenya." @barrett2001
		access to financial capital important to income diversification for fishing communities in Kenya @olale2012
		overall, diversification was positively related with total income for those living in "forest margins" @illukpitiya2008
	ifqs
		catch shares may help prevent fisheries collapse @costello2008
		entry expensive
		diversification declined at a time following catch share programs on the West Coast @holland2016
		vessels that entered West Coast fisheries less diversified than those that entered before IFQs (and of course cost is increased to get into these fisheries for those that were not already participants) @holland2016
		catch share programs may reduce variability and offset the effects of lower diversity -- found that year-to-year vessel revenue decreased after catch shares (and average revenue increased for most) @holland2016

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. We expect future research examining diversification over
finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events (e.g. the Exxon Valdez oil spill in 1989) that had significant
consequences for the ecosystem and social system in Alaska (REF). In addition to
variability, ... it is important to understand the benefits of diversification
in the context of rare and extreme disturbances which occur in marine systems
worldwide.

TODO: cut apart and recombine the following couple chunks:

Our results provide additional support for the idea that individual fishers can
reduce their income variability by diversifying their catch portfolios. However,
most individuals do not change their catch diversity substantially from year to
year [OLE: is there a supplemental figure or statistic that speaks to this
point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
high-reward quadrant in which reducing diversification increases both revenue
variability and expected revenue. Previous work has suggested that changes to
fisheries management such as limited entry or catch share programs may
inadvertently act as barrier to diversification and promote specializing
(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
specialized fishing strategies may be associated with increased revenue
variability, the benefit may be that they are also associated with increased
revenue.

Worldwide, fisheries confront the challenge of sustainable management (). While
many fisheries are well managed and productive, many are also rebuilding from
periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
in fisheries rebuilding is to modify regulations to reduce the unintended
harvest rebuilding stocks while allowing the continued harvest of healthy stocks
(need REF). Such bycatch reduction is an important goal for both ecological
(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
changes for bycatch reduction as well as other reasons gradually lead to
fisheries that target one or a few species and individual fishers wishing to
participate in multiple fisheries must own (or purchase) multiple permits. Thus
a general consequence of fisheries regulations that are shown to improve
sustainability of fish stocks and avoid overfishing - including improved
assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
al. 2008 ) - may ultimately restrict individual diversification or at least make
developing a diverse fishing portfolio more costly. Add Alaska example of
IFQs...?

>>>>

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels that fish a greater diversity of species have less
revenue variability than those that fish a lower species diversity. However, it
is unclear if these benefits extend to the level of the individual and how
year-to-year changes in diversification affect revenue and revenue variability.
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities and investigate
how individual actions along each axis affect revenue and revenue variability.
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. Our results demonstrate that the most effective
option for individuals to decrease revenue variability is to purchase permits
to participate in additional or more diverse fisheries. However, this option is
costly and often limited by regulations such as limited entry and catch share
programs. With increasing environmental variability in a changing climate, it
will be particularly important that individuals who rely on natural resources
for their livelihood have effective strategies to reduce income variability.

<!-- TODO: fit something about specialization over time into the abstract.
Thus, over the past 30 years as the Alaskan commercial fishing fleet has become
increasingly specialized, -->

@@

@@

Our results across fishing strategies are in line with previous findings
[@kasperski2013, @sethi2012b, @sethi2014] at the community and vessel levels,
but our results with respect to individual year-to-year behaviour provide new
insights. Within strategies, specializing tended to be associated with
increased revenue. This could be for a number of reasons. First, specialization
can increase efficiency [@smith1986] and individual behaviour is an important
component to the efficiency of catch [@hilborn1985]. Indeed, specialization can
also increase efficiency in other natural resource industries, such as rice
production [@abdulai2000]. Second, given the availability of "cash crops",
specializing can be highly profitable. For example, when cocoa prices have been
high, cocoa farmers in Nigeria have been highly specialized on the crop
[@mustapha1999] and specializing on valuable tobacco farming can be the optimal
strategy regardless of alternative crops [@chavez2014]. In our analyses,
fishers with available sockeye salmon, a highly valuable species, benefited the
most from specializing. However, specializing can have long-term risks if
marker, biological, or environmental conditions change. For instance, U.S.
farmers specialized on wheat, corn, or soybeans were hardest hit in the early
1980s when a financial crisis hit [@ikerd1990]. Third, a lack of specialization
can happen as a consequence of non-ideal circumstances such as poor
availability of a more valuable fish stock in a given year. For example,
downward economic shocks have been associated with increased diversity for
households relying on forests for income [@debela2012].

We also found that altering species diversity tended to be associated with
increased revenue variability. There are number of reasons by annual revenue
variability might increase with changes to individual fished species diversity.
First, changes to fishing patterns will sometimes be associated with a learning
or adaptation phase as individuals become accustomed to new locations, new
gear, or new species [REF]. Secondly, changes in diversity might reflect years
in which targeted species were less available and fishers were forced to target
other species or landed more bycatch adding stochasticity to their revenue.
Most importantly, this increased variability technically reflects an inability
for our large-scale model to predict an individual's revenue in a subsequent
year. There are many reasons, often unique to an individual permit holder's
behaviour [@hilborn1985] and therefore not included in a large-scale model, that
might make that specific individual's revenue less predictable if species
diversity changes.

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial cost. Related barriers are apparent in other
natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008]. Another major barrier to
diversification can be management restrictions. In commercial fisheries, while
catch shares [@costello2008] and limited season lengths [REF] can be effective
tools to avoid overfishing and distribute social capital, they can also limit
opportunities for diversification [@smith1986; @kasperski2013]. However, catch
shares, in particular, may also reduce income variability (Fig.
S\ref{fig:ifq}A) perhaps offsetting the effects of reduced diversification
[@holland2016].

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue
and revenue variability. We expect future research examining diversification
over finer scales will uncover additional relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [REF].

<!-- TODO: cut apart and recombine the following couple chunks:-->

<!--Our results provide additional support for the idea that individual fishers can
reduce their income variability by diversifying their catch portfolios. However,
most individuals do not change their catch diversity substantially from year to
year [OLE: is there a supplemental figure or statistic that speaks to this
point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
high-reward quadrant in which reducing diversification increases both revenue
variability and expected revenue. Previous work has suggested that changes to
fisheries management such as limited entry or catch share programs may
inadvertently act as barrier to diversification and promote specializing
(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
specialized fishing strategies may be associated with increased revenue
variability, the benefit may be that they are also associated with increased
revenue.-->

<!--
Worldwide, fisheries confront the challenge of sustainable management (). While
many fisheries are well managed and productive, many are also rebuilding from
periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
in fisheries rebuilding is to modify regulations to reduce the unintended
harvest rebuilding stocks while allowing the continued harvest of healthy stocks
(need REF). Such bycatch reduction is an important goal for both ecological
(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
changes for bycatch reduction as well as other reasons gradually lead to
fisheries that target one or a few species and individual fishers wishing to
participate in multiple fisheries must own (or purchase) multiple permits. Thus
a general consequence of fisheries regulations that are shown to improve
sustainability of fish stocks and avoid overfishing - including improved
assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
al. 2008 ) - may ultimately restrict individual diversification or at least make
developing a diverse fishing portfolio more costly. Add Alaska example of
IFQs...?
-->
